---
title: Mean, SD, Standardization
author: Felix
date: ''
slug: []
categories: []
tags: []
---

# Load in packages 
```{r}
library(tidyverse)
```

# Take a look at the data 

```{r}
str(mtcars)
```

# Calculating the mean of multiple columns   

```{r}
mtcars %>% head() %>% mutate(mean = rowMeans(select(.,vs,am))) 
```

Calculate the mean of the columns with "m." Useful when calculating the means of the questionnaire items that are named with the name of the questionnaire 
```{r}
mtcars %>% head() %>% mutate(mean = rowMeans(select(.,contains("m")))) 
```

Similar to above 
```{r}
mtcars %>% head() %>% mutate(mean = rowMeans(select(.,starts_with("c"))))
```

# Calculate the standard deviations of multiple columns 
Because there is no functions as straightforward as `colMeans`, I use the basic R function `apply`. 

```{r}
mtcars %>% head() %>% mutate(sd = apply(X = select(., starts_with("c")), FUN = sd, MARGIN = 1))
```

# Calculate the mean and sd of a variable in the sample 

```{r}
temp = mtcars %>% transmute(var1 = rowMeans(select(.,disp,hp)), 
                  var2 = rowMeans(select(.,drat,wt)))
temp %>% head()
temp %>% summarize(mean_var1 = mean(var1), sd_var1 = sd(var1), 
                   mean_var2 = mean(var2), sd_var2 = sd(var2))

#Alternatively, if the vars are not already created
mtcars %>% 
  summarize(mean_var1 = mean(rowMeans(select(.,disp,hp))),
            sd_var1 = sd(rowMeans(select(.,disp,hp))),
            mean_var2 = mean(rowMeans(select(.,drat,wt))),
            sd_var2 = sd(rowMeans(select(.,drat,wt))),)
```

# Standardize variable(s)

```{r}
#Standardize one variable 
temp %>% mutate(z_var1 = scale(var1, scale = T)) %>% head()

#Standardize multiple variable 
temp %>% mutate(across(c(var1, var2), #select vars 
                       scale, scale = T, #apply function 
                       .names = "z_{.col}")) %>%  #name the vars (if default, old vars are replaced)
    mutate(across(everything(), round, 3)) %>% head()
```


# Resources:

- Documentation of `across`: https://dplyr.tidyverse.org/reference/across.html


-----------------
_Disclaimer: These codes are not necessarily the best (e.g., fastest) solutions to the tasks, but simply the codes I use in my own work. I would deeply appreciate any suggestions to optimize them. Please feel free to reach out to me via email!_
